\documentclass[a4paper,12pt]{article}

\input{../../preamble.tex}
\graphicspath{{../figs/}}

%--------------------------
% PACKAGES
%--------------------------
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{geometry}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage[normalem]{ulem}

%--------------------------
% PAGE SETUP
%--------------------------
\geometry{margin=1in}
\pagestyle{fancy}
\fancyhf{}
\rhead{\thepage}
\lhead{Learning Journal}

%--------------------------
% SECTION FORMATTING
%--------------------------
\titleformat{\section}[block]{\large\bfseries}{Session \thesection:}{1em}{}
\titleformat{\subsection}[block]{\bfseries}{\thesubsection}{1em}{}
\setcounter{secnumdepth}{2}

%--------------------------
% CUSTOM ENVIRONMENTS
%--------------------------
\newenvironment{deff}[1][Definition]{\par\vspace{1em}\noindent\textbf{#1.}\hspace{0.5em}}{\par\vspace{0.1em}}

\newenvironment{rem}[1][Remark]{\par\vspace{1em}\noindent\textit{#1.}\hspace{0.5em}\begingroup\itshape}{\endgroup\par\vspace{0.1em}}

% Explanation of the custom environment:
% 1. \newenvironment{definition}[1][Definition]: This creates a new environment named "definition" with an optional argument. The default title is "Definition."
% 2. \par\vspace{1em}: Adds vertical space before the environment begins.
% 3. \noindent\textbf{#1.}: Formats the title (e.g., "Definition.") as bold text.
% 4. \hspace{0.5em}: Adds a small space between the title and the definition text.
% 5. The closing curly braces define the formatting after the environment ends, including vertical spacing.


%--------------------------
% DOCUMENT
%--------------------------
\begin{document}

\begin{center}
    {\Huge \textbf{Learning Journal}} \\
    \vspace{0.5em}
    {\large Organized by Sessions with Topic Summaries}
\end{center}
\vspace{1em}
\hrule
\vspace{2em}

%--------------------------
% TEMPLATE STRUCTURE
%--------------------------
\section*{Instructions}
\begin{itemize}
    \item Use this document to track your learning sessions.
    \item Each session entry should include: \begin{enumerate}
        \item Topics covered (briefly).
        \item Key insights or definitions.
        \item Problems attempted (with references or solutions if necessary).
        \item Questions or areas for follow-up.
    \end{enumerate}
    \textit{Make Anki flashcards of whatever makes sense. Even topics if you'd like, just to leverage their spaced repitition algorithm.}
    \item After completing a topic, write a summary essay, incorporating insights from session notes and solved examples.
\end{itemize}
\newpage

%--------------------------
% RESOURCES AND LEARNING PLANS
%--------------------------
\input{../resources.tex}
\newpage

\section*{Learning Plan}
\textit{This is a first pass on some topics I find interesting. I don't need to understand everything the first time. I don't need to go through all the subtopics the first time. I need to understand the main idea, understand why and when it is useful, and try my hand at some basic problems to use it myself. I can always pick up these topics again since I keep a record here. Just remember to jot down your questions, the things you don't understand (and why) and what you'd like to investigate further in the future. You can \textbf{always} come back!}
\input{../learning-plans/first-pass.tex}
\newpage

%--------------------------
% SESSIONS
%--------------------------
\input{sessions/03-01-25.tex}
\newpage
\input{sessions/04-01-25.tex}
\newpage
\input{sessions/05-01-25.tex}
\newpage
\input{sessions/06-01-25.tex}
\newpage
\input{sessions/08-01-25.tex}
\newpage
\input{sessions/09-01-25.tex}


\end{document}
