\subsection*{Calculus of Variations in Physics (second pass)}   
\begin{itemize}
    \item Michael Penn Video as a recap
    \item Derive the geodesic on a sphere \textred{recap!}
    \item Ask GPT for some interesting physics problems where one sees how powerful the formulation is powers. Try to derive steps yourself in these examples.
    \item The relation to Hamiltonian mechanics
    \item Answer more questions below and synthesize and improve notes on the topic
\end{itemize}
\textbf{Questions}
\begin{itemize}
    \item How is it used in modern physics today?
    \item What are some "famous" action "derivations"?
    \item Does one use different "actions" when working on separate problems. If so, could one find a problem "midway" between those problems and see what the action looks like there? Maybe smoothly interpolate an action between these two problems to gain a deeper understanding of how and why they need different descriptions. Maybe find a general description which they are both special cases of. 
\end{itemize} 
\subsection*{Special Relativity, Classical Field Theory and Tensors (second pass)}
\begin{itemize}
    \item Understand the \(D\)-dimensional scalar field derivation on the on-shell wikipedia-page of the stress-energy tensor
    \item Solving some harder problems with 4-vectors
    \item Revisit Goldstein's 4-vector Lagrangian Field section and understand the derivation of Nöether's theorem.
    \item Synthesize and improve notes
\end{itemize}
\subsection*{Complex Analysis in Quantum Mechanics}
\subsection*{Complex Analysis in Quantum Field Theory and relations to Lie Theory}
\subsection*{MIT Quantum Mechanics II Course}